---
title: 'Let chatGPT collect more data'
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
# This chunk is for initial setup
knitr::opts_chunk$set(echo = TRUE)

# Function to check and install packages
check_install_package <- function(pkg) {
  if (!require(pkg, character.only = TRUE)) {
    install.packages(pkg)
    library(pkg, character.only = TRUE)
  }
}

# Check and install necessary packages
check_install_package("readr")
check_install_package("dplyr")
```

We start by loading the data.

```{r}
# Import the listings.csv file
listings <- read_csv("../dataRaw/listings.csv", show_col_types = FALSE)
```

We will ask chatGPT to gather some data. For this, we first need to export a list with the names of all neighborhoods.

```{r}
# Extract unique neighborhood names
unique_neighbourhoods <- unique(listings$neighbourhood_cleansed)

# Write to a CSV file
write.csv(unique_neighbourhoods, "unique_neighbourhoods.csv", row.names = FALSE)
```

Then we ask chatGPT with a prompt like this to collect some information:

  I have a CSV file containing the names of neighborhoods in the city of Rotterdam. I’d like you to gather detailed information for each neighborhood and assign ratings (on a scale from 0 to 10) for the following variables:
  
  Neighborhood Coolness: How attractive the neighborhood is to people aged 20-30.
  Neighborhood Centrality: How centrally located the neighborhood is within the city.
  Neighborhood Quietness: How quiet the neighborhood is at night.
  Neighborhood Fanciness: The level of sophistication and quality of shops in the area.
  Please provide unique ratings for each neighborhood based on research, avoiding placeholder values.
  
  Please return results in a markdown file.

The result is in the file `complete_rotterdam_neighborhood_ratings.csv`.
