---
title: 'Chapter 2: simple regression'
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
# This chunk is for initial setup
knitr::opts_chunk$set(echo = TRUE)

# Function to check and install packages
check_install_package <- function(pkg) {
  if (!require(pkg, character.only = TRUE)) {
    install.packages(pkg)
    library(pkg, character.only = TRUE)
  }
}

# Check and install necessary packages
check_install_package("stargazer")
check_install_package("dplyr")

# Load necessary libraries
library(ggplot2)
library(dplyr)  # for data manipulation
```

We start again by loading the data.

```{r}
# Load the datasets from the RData file
load("../dataCreated/listings_clean.RData")

# Display the first few rows of the cleaned dataset
head(listings_clean)
```

Filter the data so that we use only listings for at most 6 people. From now on use this

```{r}
listings_clean_filtered <- listings_clean %>%
  filter(accommodates <= 6)
```

Regress price on `review_scores_rating`.

```{r}
# Run a linear regression of price on review_scores_rating
model <- lm(price ~ review_scores_rating, data = listings_clean_filtered)

# Display the summary of the regression model
summary(model)
```

Run a regression of price on how many people can be accommodated.

```{r}
# Run a linear regression of price on accommodates
model_accommodates <- lm(price ~ accommodates, data = listings_clean_filtered)

# Display the summary of the regression model
summary(model_accommodates)
```

On slide 21, we have said that the sample moment conditions immediately imply that:
1. the average residual is zero
2. the average covariance between the residual and the regressor is zero
3. the regression line goes through the average of the dependent variable and the regressor.

We now verify this.

```{r}
# 1. Verify that the average residual is zero
residuals_accommodates <- residuals(model_accommodates)
mean_residuals <- mean(residuals_accommodates)
cat("1. Average of residuals:", mean_residuals, "\n")

# 2. Verify that the average covariance between the residual and the regressor (accommodates) is zero
cov_residuals_regressor <- cov(residuals_accommodates, listings_clean_filtered$accommodates)
cat("2. Covariance between residuals and accommodates:", cov_residuals_regressor, "\n")

# 3. Verify that the regression line goes through the averages of the dependent variable and the regressor
mean_price <- mean(listings_clean_filtered$price)
mean_accommodates <- mean(listings_clean_filtered$accommodates)

# Calculate the predicted price at the average of accommodates
predicted_price_at_mean <- coef(model_accommodates)[1] + coef(model_accommodates)[2] * mean_accommodates

cat("3. Average price:", mean_price, "\n")
cat("   Predicted price at average accommodates:", predicted_price_at_mean, "\n")
```

We have also said that the total sum of squares for the dependent variable is equal to the explained sum of squares plus the residual sum of squares. We can also verify this.

```{r}
# 1. Calculate the Total Sum of Squares (TSS)
mean_price <- mean(listings_clean_filtered$price)
TSS <- sum((listings_clean_filtered$price - mean_price)^2)
cat("Total Sum of Squares (TSS):", TSS, "\n")

# 2. Calculate the Residual Sum of Squares (RSS)
RSS <- sum(residuals_accommodates^2)
cat("Residual Sum of Squares (RSS):", RSS, "\n")

# 3. Calculate the Explained Sum of Squares (ESS)
# ESS is the difference between TSS and RSS
ESS <- TSS - RSS
cat("Explained Sum of Squares (ESS):", ESS, "\n")

# 4. Verify that TSS = ESS + RSS
cat("TSS == ESS + RSS:", TSS == (ESS + RSS), "\n")
```

From this we can also compute the $R^2$ measure by hand.

```{r}
# 1. Calculate Total Sum of Squares (TSS)
mean_price <- mean(listings_clean_filtered$price)
TSS <- sum((listings_clean_filtered$price - mean_price)^2)

# 2. Calculate Residual Sum of Squares (RSS)
RSS <- sum(residuals_accommodates^2)

# 3. Calculate Explained Sum of Squares (ESS)
ESS <- TSS - RSS

# 4. Calculate R^2 using the formula R^2 = ESS / TSS or 1 - RSS / TSS
R_squared <- 1 - (RSS / TSS)

# Display the result
cat("R^2 (by hand):", R_squared, "\n")
```

It is the same as above in the regressian output (which of course has to be the case)!

Next we use the Stargazer package that we loaded in the beginning to produce a nice table with regression results for three different specifications. See slide 24 for details.

```{r}
# 1. Run a level-level regression of price on accommodates
level_level <- lm(price ~ accommodates, data = listings_clean_filtered)

# 2. Run a log-level regression of log(price) on accommodates
log_level <- lm(log(price) ~ accommodates, data = listings_clean_filtered)

# 3. Run a log-log regression of log(price) on log(accommodates)
log_log <- lm(log(price) ~ log(accommodates), data = listings_clean_filtered)

# 4. Create a table with stargazer displaying all three results
stargazer(level_level, log_level, log_log,
          type = "text",
          title = "Regression Results",
          column.labels = c("Level-Level", "Log-Level", "Log-Log"),
          dep.var.labels = "Price",
          covariate.labels = c("Accommodates", "Constant"),
          omit.stat = c("f", "ser"),
          digits = 3)
```

Finally, we produce a figure with the data points and the three fitted regression lines.

```{r}
# 1. Create a new data frame with predictions
listings_clean_filtered$pred_level_level <- predict(level_level, newdata = listings_clean_filtered)
listings_clean_filtered$pred_log_level <- exp(predict(log_level, newdata = listings_clean_filtered))
listings_clean_filtered$pred_log_log <- exp(predict(log_log, newdata = listings_clean_filtered))

# 2. Calculate the average price for each value of accommodates
avg_prices <- listings_clean_filtered %>%
  group_by(accommodates) %>%
  summarise(avg_price = mean(price))

# 3. Scatter plot of the actual data (price vs accommodates)
p <- ggplot(listings_clean_filtered, aes(x = accommodates, y = price)) +
  geom_point(alpha = 0.5) +
  labs(title = "Estimated Empirical Relationships",
       x = "Accommodates",
       y = "Price") +
  theme_minimal()

# 4. Add the fitted values from the level-level model
p <- p + geom_line(aes(y = pred_level_level, color = "Level-Level"), linewidth = 1)

# 5. Add the fitted values from the log-level model
p <- p + geom_line(aes(y = pred_log_level, color = "Log-Level"), size = 1)

# 6. Add the fitted values from the log-log model
p <- p + geom_line(aes(y = pred_log_log, color = "Log-Log"), size = 1)

# 7. Add the average price for each accommodates value as red squares
p <- p + geom_point(data = avg_prices, aes(x = accommodates, y = avg_price), 
                    color = "red", shape = 15, size = 3)

# 8. Add legend and show the plot
p <- p + scale_color_manual(name = "Model",
                            values = c("Level-Level" = "blue", "Log-Level" = "green", "Log-Log" = "red"))

# Display the plot
print(p)

```



