---
title: 'Chapter 1: open data and perform descriptive analysis'
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
# This chunk is for initial setup
knitr::opts_chunk$set(echo = TRUE)

# Function to check and install packages
check_install_package <- function(pkg) {
  if (!require(pkg, character.only = TRUE)) {
    install.packages(pkg)
    library(pkg, character.only = TRUE)
  }
}

# Check and install necessary packages
check_install_package("ggplot2")
check_install_package("dplyr")
```

We start by loading the data. Recall that we created `listings_clean` and `listings_clean_accommodates_2`. The former selects observations with finite prices and review scores, more than 10 received reviews, and a price below 300 USD; the latter selects listings that accommodate 2 people.

```{r}
# Load data
# Load the datasets from the RData file
load("../dataCreated/listings_clean.RData")

# Display the first few rows of the cleaned dataset
head(listings_clean)
```

Create some summary statistics.

```{r}
# Create a summary statistics table with one row per variable
summary_stats <- tibble::tibble(
  Variable = c("Price", "Review Scores Rating", "Accommodates"),
  Mean = c(round(mean(listings_clean$price, na.rm = TRUE), 2),
           round(mean(listings_clean$review_scores_rating, na.rm = TRUE), 2),
           round(mean(listings_clean$accommodates, na.rm = TRUE), 2)),
  Median = c(round(median(listings_clean$price, na.rm = TRUE), 2),
             round(median(listings_clean$review_scores_rating, na.rm = TRUE), 2),
             round(median(listings_clean$accommodates, na.rm = TRUE), 2)),
  Standard_Deviation = c(round(sd(listings_clean$price, na.rm = TRUE), 2),
                         round(sd(listings_clean$review_scores_rating, na.rm = TRUE), 2),
                         round(sd(listings_clean$accommodates, na.rm = TRUE), 2)),
  Min = c(round(min(listings_clean$price, na.rm = TRUE), 2),
          round(min(listings_clean$review_scores_rating, na.rm = TRUE), 2),
          round(min(listings_clean$accommodates, na.rm = TRUE), 2)),
  Max = c(round(max(listings_clean$price, na.rm = TRUE), 2),
          round(max(listings_clean$review_scores_rating, na.rm = TRUE), 2),
          round(max(listings_clean$accommodates, na.rm = TRUE), 2))
)

# Display the summary statistics table
summary_stats
```

Histogram of `price` for those apartments accommodating 2 people.

```{r}

# Create the histogram
ggplot(listings_clean_accommodates_2, aes(x = price)) +
  geom_histogram(binwidth = 10, fill = "blue", color = "black", alpha = 0.7) +
  labs(
    title = "Histogram of Price for Listings that Accommodate 2 People",
    x = "Price",
    y = "Count"
  ) +
  theme_minimal()
```

What is the empirical relationship between the price and the review score for an apartment that accommodates 2 people? One can look at a scatter plot.

```{r}
# Create the scatter plot
ggplot(listings_clean_accommodates_2, aes(x = review_scores_rating, y = price)) +
  geom_point(alpha = 0.6) +
  labs(
    title = "Scatter Plot of Price vs. Review Scores Rating (Accommodates 2)",
    x = "Review Scores Rating",
    y = "Price"
  ) +
  theme_minimal()
```

Or at a correlation.

```{r}
# Calculate the correlation between price and review_scores_rating
correlation <- cor(listings_clean_accommodates_2$price, listings_clean_accommodates_2$review_scores_rating, use = "complete.obs")

# Display the correlation
correlation
```

How does the price for an apartment that accommodates 2 vary by neighborhood?

```{r}
# Create a summary table with the average price for each neighborhood
neighborhood_price_summary <- listings_clean_accommodates_2 %>%
  group_by(neighbourhood) %>%
  summarise(average_price = round(mean(price, na.rm = TRUE), 2)) %>%
  arrange(neighbourhood)

# Display the summary table
neighborhood_price_summary
```

How does the price vary with the number of people that can stay in an apartment. We wille explore this using a set of so-called box plots.

A box plot, or box-and-whisker plot, visually displays the distribution of data using five key metrics: the minimum, first quartile (Q1), median (Q2), third quartile (Q3), and maximum. The box shows the interquartile range (IQR), representing the middle 50% of the data, with the line inside the box indicating the median. Whiskers extend from the box to the smallest and largest values within 1.5 times the IQR. Outliers, or points outside this range, are shown as individual dots. Box plots are useful for quickly assessing the spread, center, and potential outliers in the data.

We make a figure with one box plot for each value of `accommodates`.

```{r}
# Create box plots of price by accommodates
ggplot(listings_clean, aes(x = factor(accommodates), y = price)) +
  geom_boxplot() +
  labs(
    title = "Box Plot of Price by Accommodates",
    x = "Number of People Accommodated",
    y = "Price"
  ) +
  theme_minimal()
```

